JEThread: A Thread Pool Implementation in C

### 2018-05-04 thread-pool-implementation/paper.tex

In addition to batch assigning jobs, we also treat blocking and nonblocking tasks differently. This is necessary because JEThread by default starts with $N$ threads, where $N$ is the number of processors, and if all workers threads are executing some blocking tasks, then nonblocking tasks will not be able to execute until much later. It becomes even worse when majority of tasks are blocking tasks. In such case, the fixed number of workers would be blocked constantly. To solve this problem, we use an approach similar to Golang's parking threads, where we create a new thread to handle blocking tasks and only use our default set of worker threads to handle nonblocking tasks. In order to implement this, we need to identify if a task is blocking or nonblocking. This is difficult and inefficient to do at runtime, but can be done much easier in compile time. In our implementation, we rely on user to specify the CallType when adding a task. This is not fundamentally necessary - we could have modified GCC compiler or inspected task at preprocessing time to accomplish the same thing, but due to the time constraint we exposed this option to users. We believe that users would try their best to provide accurate CallType because they want better performance. Even if they accidentally misidentified CallType, this would not effect the correctness of JEThread.

>>>>

\subsection{Blocking versus Non-Blocking Tasks}
In addition to batch assigning jobs, we also treat blocking and nonblocking tasks differently. This is necessary because JEThread by default starts with $N$ threads, where $N$ is the number of processors, and if all workers threads are executing some blocking tasks, then nonblocking tasks will not be able to execute until much later. It becomes even worse when majority of tasks are blocking tasks. In such case, the fixed number of workers would be blocked constantly. To solve this problem, we use an approach similar to Golang's parking threads, where we create a new thread to handle blocking tasks and only use our default set of worker threads to handle nonblocking tasks. In order to implement this, we need to identify if a task is blocking or nonblocking. This is difficult and inefficient to do at runtime, but can be done much easier in compile time. In our implementation, we rely on user to specify the CallType when adding a task. This is not fundamentally necessary - we could have modified GCC compiler or inspected task at preprocessing time to accomplish the same thing, but due to the time constraint we exposed this option to users. We believe that users would try their best to provide accurate CallType because they want better performance. Even if they accidentally misidentified CallType, this would not effect the correctness of JEThread.

\subsection{Work Stealing}
One of our goals is to achieve automatic load balancing. For this purpose, we implement work stealing for each worker thread. For each worker, we first do a sem\_trywait() to see if there is any job on its private queue. If there is none, then it randomly selects a victim thread and tries to steal a task from the victim and executes it. If keeps doing so until there is no more work to steal or if it has stolen $N$ times, where $N$ is the number of workers. This is our first attempt at work stealing. To optimize this, we could have kept randomly selecting victims until there is a victim with work above a certain threshold or up to a certain number of tries. To avoid all workers trying to steal when there is no tasks available, we can also use randomization to determine if a worker should try work stealing or just block until it is assigned a job. Another optimization we can perform is to steal more than one work at a time since we want it to be worth the overhead of running the algorithm. Although not reflected in the evaluation, we did try to evaluate the impact of the current work stealing techniques and found that even the naive approach improves JEThread's performance slightly.
